\section{Introduction}
\subsection{Prior work}
%TODO: refer to new whitepaper when launched
Carbon dating, as described by Clark\cite{clark2012commitcoin} was using bitcoin to commit to a message with a provable timestamp by embedding a commitment in a bitcoin block. Todd\cite{todd2016opentimestamps} introduced the first publicly available timestamp service with Open Timestamps in 2016. It enables committing to a state of a document by adding the hash of it to a Merkle tree that is then written into a Bitcoin transaction. By producing a proof of inclusion in the Merkle tree and a proof that the root hash is included in a given block in the Bitcoin blockchain any third party can verify that the document had the given state at that point in time. Centrifuge uses smart contracts on the Ethereum blockchain to timestamp the documents state in a very similar way.

Coleman's work on state channels\cite{coleman2018counterfactual}, the generalized version of a payment channel first introduced by Poon et al.\cite{poon2016lightningnetwork} explores many of the challenges associated with settling peer to peer transactions. Channels must first be opened on-chain, but can then be modified very cheaply by sending signed messages back and forth. Expensive on-chain transactions are only used in case of disagreement, or to close a channel. Many of the attacks on state channel protocols were considerations in the design of the Centrifuge protocol. While Centrifuge does not strictly adhere to the definition of a state channel, it addresses many problems similarly.

Quorum's private transaction layer\cite{quorumprivatetransactionwiki} describes a mechanism by which the state of a contract is encrypted and only updated and validated by the parties that have access to the smart contract. A transaction on-chain is used to update the hash of the off-chain state, but doesn't give any guarantees that the transaction was valid.
\subsection{Current State}
This paper defines the current implementation of the Centrifuge protocol. It represents a work in progress and will be updated to reflect the newest version of the protocol deployed. Centrifuge launched in early 2019 with the version specified for the first time here. The current version should be viewed as a first minimum viable product. Refer to the section on notable revisions (ref. \ref{sec:revisions}) for a list of prior versions of this document.
